{\boldmath Pseudoscalar top-Higgs coupling: Exploration of $\mathrm{CP}$-odd observables to resolve the sign ambiguity }

### 2016-03-04 Paper_ttbarH_Draft.tex

For comparison purposes, we have also used our simulated events
to test the asymmetry corresponding to the lab frame observable given
in Ref.~\cite{Guadagnoli}. {\bf (KK: Is this lab frame observable
  just a TP, like we have used?  It seems strange to talk about an
  asymmetry associated with an observable, when normally an asymmetry
  is itself considered to be an observable.  I'm wondering if we are
  considering an asymmetry associated with a TP, or if we are actually
  just looking at whatever observable they have defined\ldots.  Sorry for
  my confusion on this!)}  We have found that this observable
{\bf (KK: ``asymmetry''?)}
appears to
be slightly less sensitive than the combination $\epsilon_{10}$,
{\bf (KK: ``$\mathcal{A}(\epsilon_{10})$''?)}
giving rise to a separation between the $\mathrm{CP}$-mixed scenarios
that is smaller by about $1.4\sigma$.

@@

In our numerical analyses so far we have used relatively large
samples of events ($10^5$ events per sample)
in order to clearly distinguish
which observables would be most promising.  
The number of events expected at the HL-LHC, however, is
smaller than the number of events that we have used in our
simulations.  In this section we reexamine
the more promising observables, using sample sizes that are
more attainable in the near future.  {\bf (KK: OK?)}

Let us first make some estimates regarding the number of signal events
expected at the HL-LHC.
In Sec.~\ref{sec3} we introduced several mild selection cuts.
Implementing these cuts, and assuming that the final state leptons
could be either electrons or muons, {\bf (KK: factor of 4\ldots?)}
the SM cross section for $\ppprocess$ at
$14\,\mathrm{TeV}$ is $\sim 15.3\,\mathrm{fb}$; thus, the number of
events expected within the context of the HL-LHC is $\sim
15.3\,\mathrm{fb} \times 3000\,\mathrm{fb}^{-1} = 4.59\times
10^4$. This number is expected to be larger
if $\kpt\neq 0$ (assuming $\kp=1$),
since the corresponding cross section is
larger than the SM cross section in this case.
Taking into account NLO corrections (to the
production process) via a $K$ factor of approximately
$1.2$~\cite{Dawson,Beenakker,Dittmaier}, we find that
the expected number of
events increases to $\sim 5.49 \times 10^4$.
{\bf (KK: should we use 5.5 at this point, since
  the $K$ factor is only two sig figs?)}
On the other hand, additional cuts, as
well as a reduction in efficiency related to momentum reconstruction,
will lead to a decrease in this number.  In
order to measure the asymmetry $\mathcal{A}(\epsilon_4)$,
for example, the $t$ and $\tbar$
momenta need to be reconstructed. This is challenging, not only due to
the presence of two neutrinos in the final state (which escape the
detector undetected), but also because final state objects need to be
associated with their respective parent quark~\cite{atlasconf}.
{\bf (KK: maybe we can talk about this for a minute; I don't think I
  understand it\ldots.  Wait -- do we mean that the quarks and leptons
  need to be associated with either the top or antitop?  If so, maybe
  we can reword this a bit.  Suggestion follows\ldots, maybe this
  is too explicit now, because we also wanted it to refer to the Higgs.)}
but also because the (visible) quarks and charged leptons in the final
state need to be correctly associated with the corresponding
parent particle (i.e., the top or antitop quark)~\cite{atlasconf}.
As was
already noted in Sec.~\ref{sec5}, one possibility is to use the
neutrino weighting technique along with the kinematic equations
arising from kinematic constraints related to the top and $W$ masses
as well as from the energy-momentum conservation at each of the decay
vertices involved in the process. {\bf (KK: Can we either simplify the
  previous sentence or split it
  into 2 or 3 sentences?  I'm not 100\% clear on which ideas go
  together, so I wonder if one of you could give it a try.  My preference
  would be to simplify it to something like ``\ldots weighting technique,
  along with associated kinematic constraints'' or something like
  that, and then give a reference.)}
Within the context of $t\tbar$ production {\bf (KK: correct?)} this
procedure has been used, for instance, to obtain 
spin correlation~\cite{atlasconf} and charge
asymmetry~\cite{atlascharge} measurements.
Also, events reconstructed using this technique have
been used in the analysis of angular distributions that
are useful for discriminating the signal from the backgrounds in
$t\tbar H (H\rightarrow b\bbar)$~\cite{dosSantos}.
{\bf (KK: do we need an initial state
  on this process?)} In all these cases the corresponding
efficiency in the reconstruction of the momenta is up to $\sim 80\%$.
{\bf (KK: Does this mean that none of them exceeds 80\% or that
  they all have an efficiency of at least 80\%?)}

Given the discussion in the previous paragraph, we
have generated sets of $5\times 10^4, 1\times 10^4$ and $5\times 10^3$
events and have recalculated the most sensitive observable, 
$\mathcal{A}(\epsilon_4)$, for each case. The results are displayed in
Table \ref{table11}, where it can be seen that for $5\times 10^4$ events
(which is close to our rough estimate above for the
total number of signal events for the
HL-LHC), {\bf (KK: stated correctly?)}
the observable is still very sensitive to $\kpt$.
In this case, the $\mathrm{CP}$-mixed scenarios are
effectively separated by $19\sigma$. As expected, the
sensitivity worsens as the number of events is reduced, but even with
$5\times 10^3$ events the separation between the $\mathrm{CP}$-mixed
scenarios under consideration is $\sim 6.5\sigma$. \par

@@

Finally, it is important to mention that a realistic
analysis of the sensistivity of the observables discussed in this
paper requires a study of the impact of the backgrounds, as well as
the hadronization of the quarks in the final state and the effects of
the detector. If we consider the dominant decay mode of the Higgs boson,
$H\rightarrow b\bbar$, in order to maximize the cross section of the
process, the signature is given by $4$ $b$-jets, two leptons and
missing energy.  The main background arises from the production of
$t\tbar$ in association with additional jets, with the dominant
source being the production of $t\tbar + b\bbar$.
In Ref.~\cite{chinos} it is shown that the application of a small
set of cuts results in a large improvement in the signal to background
ratio.  {\bf (KK: please check that I didn't say this incorrectly.)}
On the experimental side, a
rigorous treatment of the signal and backgrounds
for the process {\bf (KK: \ldots process\ldots)}
is performed in Ref.~\cite{atlasger}, using $20.3\,
\mathrm{fb}^{-1}$ of data at $\sqrt{s}=8\,\mathrm{TeV}$.
       {\bf (KK: please check that I didn't alter the meaning
         of this last sentence when I edited it.  Also, do we need
         to say anything about this study other than that it was done?)}

>>>>

For comparison purposes, we have also used our simulated events
to test the lab frame observable given
in Ref.~\cite{Guadagnoli}. {\bf (KK: Is this lab frame observable
  just a TP, like we have used?  It seems strange to talk about an
  asymmetry associated with an observable, when normally an asymmetry
  is itself considered to be an observable.  I'm wondering if we are
  considering an asymmetry associated with a TP, or if we are actually
  just looking at whatever observable they have defined\ldots.  Sorry for
  my confusion on this!) (N: They compute the asymmetry of a quantity that is basically the product
  of an angle and the sign of a TP)}  We have found that this observable
appears to
be slightly less sensitive than $\mathcal{A}(\epsilon_{10})$,
giving rise to a separation between the $\mathrm{CP}$-mixed scenarios
that is smaller by about $1.4\sigma$.

@@

In our numerical analyses so far we have used relatively large
samples of events ($10^5$ events per sample)
in order to clearly distinguish
which observables would be most promising.  
The number of events expected at the HL-LHC, however, is
smaller than the number of events that we have used in our
simulations.  In this section we reexamine
the more promising observables, using sample sizes that are
more attainable in the near future.

Let us first make some estimates regarding the number of signal events
expected at the HL-LHC.
In Sec.~\ref{sec3} we introduced several mild selection cuts.
Implementing these cuts, and assuming that the final state leptons
could be either electrons or muons, {\bf (KK: factor of 4\ldots?)}
the SM cross section for $\ppprocess$ at
$14\,\mathrm{TeV}$ is $\sim 15.3\,\mathrm{fb}$; thus, the number of
events expected within the context of the HL-LHC is $\sim
15.3\,\mathrm{fb} \times 3000\,\mathrm{fb}^{-1} = 4.59\times
10^4$. This number is expected to be larger
if $\kpt\neq 0$ (assuming $\kp=1$),
since the corresponding cross section is
larger than the SM cross section in this case.
Taking into account NLO corrections (to the
production process) via a $K$ factor of approximately
$1.2$~\cite{Dawson,Beenakker,Dittmaier}, we find that
the expected number of
events increases to $\sim 5.5 \times 10^4$.
On the other hand, additional cuts, as
well as a reduction in efficiency related to momentum reconstruction,
will lead to a decrease in this number.  In
order to measure the asymmetry $\mathcal{A}(\epsilon_4)$,
for example, the $t$ and $\tbar$
momenta need to be reconstructed. This is challenging, not only due to
the presence of two neutrinos in the final state (which escape the
detector undetected), but also because the (visible) quarks and charged leptons in the final
state need to be correctly associated with the corresponding
parent particle (i.e., the top or antitop quark)~\cite{atlasconf}.
As was
already noted in Sec.~\ref{sec5}, one possibility is to use the
neutrino weighting technique along with associated kinematic constraints {\bf (N:add references)}.
Within the context of $t\tbar$ production this
procedure has been used, for instance, to obtain 
spin correlation~\cite{atlasconf} and charge
asymmetry~\cite{atlascharge} measurements.
Also, events reconstructed using this technique have
been used in the analysis of angular distributions that
are useful for discriminating the signal from the backgrounds in
$t\tbar H (H\rightarrow b\bbar)$ at the LHC~\cite{dosSantos}.
 In all these cases the corresponding
efficiency in the reconstruction of the momenta is up to $\sim 80\%$.
{\bf (KK: Does this mean that none of them exceeds 80\% or that
  they all have an efficiency of at least 80\%?)}

Given the discussion in the previous paragraph, we
have generated sets of $5\times 10^4, 1\times 10^4$ and $5\times 10^3$
events and have recalculated the most sensitive observable, 
$\mathcal{A}(\epsilon_4)$, for each case. The results are displayed in
Table \ref{table11}, where it can be seen that for $5\times 10^4$ events
(which is close to our rough estimate above for the
total number of signal events for the
HL-LHC), 
the observable is still very sensitive to $\kpt$.
In this case, the $\mathrm{CP}$-mixed scenarios are
effectively separated by $19\sigma$. As expected, the
sensitivity worsens as the number of events is reduced, but even with
$5\times 10^3$ events the separation between the $\mathrm{CP}$-mixed
scenarios under consideration is $\sim 6.5\sigma$. \par

@@

Finally, it is important to mention that a realistic
analysis of the sensistivity of the observables discussed in this
paper requires a study of the impact of the backgrounds, as well as
the hadronization of the quarks in the final state and the effects of
the detector. If we consider the dominant decay mode of the Higgs boson,
$H\rightarrow b\bbar$, in order to maximize the cross section of the
process, the signature is given by $4$ $b$-jets, two leptons and
missing energy.  The main background arises from the production of
$t\tbar$ in association with additional jets, with the dominant
source being the production of $t\tbar + b\bbar$.
In Ref.~\cite{chinos} it is shown that the application of a small
set of cuts results in a large improvement in the signal to background
ratio.  
On the experimental side, a
rigorous treatment of the signal and backgrounds
for $t\tbar H$ production with $H\to b\bbar$ 
is performed in Ref.~\cite{atlasger}, using $20.3\,
\mathrm{fb}^{-1}$ of data at $\sqrt{s}=8\,\mathrm{TeV}$.
       {\bf (KK: please check that I didn't alter the meaning
         of this last sentence when I edited it.  Also, do we need
         to say anything about this study other than that it was done?)}
